Towards fast and optimal centroiding of faint stars

### 2014-12-11 ms/figure.tex
update intro

\documentclass[12pt, preprint]{aastex}
\usepackage{graphicx}	% For figures
\usepackage{natbib}	% For citep and citep
\usepackage{amsmath}	% for \iint
\usepackage{bbm}
\usepackage[breaklinks]{hyperref}	% for blackboard bold numbers
\usepackage{hyperref}
\hypersetup{colorlinks}
\usepackage{color}
\usepackage{morefloats}
\definecolor{darkred}{rgb}{0.5,0,0}
\definecolor{darkgreen}{rgb}{0,0.5,0}
\definecolor{darkblue}{rgb}{0,0,0.5}
\hypersetup{ colorlinks,
linkcolor=darkblue,
filecolor=darkgreen,
urlcolor=darkred,
citecolor=darkblue }
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}

\begin{document}
\author{
  Mohammadjavad~Vakili\altaffilmark{1},
  David~W.~Hogg\altaffilmark{1,2,3}
\altaffiltext{1}{Center for Cosmology and Particle Physics, New York University}
\altaffiltext{2}{Center for Data Science, New York University}
\altaffiltext{3}{Max-Planck-Institut f\"ur Astronomie}
}

\title{Towards fast and optimal centroiding of stars}

@@

We present a modification of the 3$\times$3 polynomial method for estimating the centroid 
of stars in astronomical images. After simulating a large number of stars with relatively low 
signal-to-noise-ratio, and in a wide range of full width at half maxima, we compare the results of our method
to those obtained from the quartic approximation method used in SDSS photometric pipeline, 
as well as finding the centroid by fitting the exact PSF model which was used to generate the 
simulated data. Comparison of the root-mean-squared error in centroid measurement, resulting 
from these techniques to the Cramer-Rao lower bound of centroid estimates, 
suggests that, the modified 3$\times$3 polynomial trick, though not as accurate as PSF fitting method
, is able to deliver sufficiently accurate estimates comparable to those obtained from the quartic approximation.

@@

A common practice in astronomy is taking imaging data, and then finding the coordinates
of various light sources across the sky. Finding unbiased estimate of the center of point
sources, convolved with telescope point spread function (and atmospheric PSF in case of
ground based telescopes), and pixel response function, is crucial to further steps of
astronomical image processing. For instance, proper measurement of the shapes of galaxies
requires interpolating the PSF from the position of sparsely located stars across the
image to the positions of galaxies. The accuracy of the PSF estimation therefore,
relies on how accurately we know the centroid of stars.

@@

Reaching the optimal measurement of the centroid of stars however, is limited
by the lack of knowledge about the exact shape of the PSF, and also presence of noise;
both sky noise and CCD readout noise. Thus, of particular interest is devising a fast method that returns
more accurate estimate of centroids in a realistic range of signal-to-noise ratios.       
To date, a number of softwares have been designed for the purpose of extracting astronomical
sources and making catalogs. One of these softwares is SExtractor \citep{sex},
whose centroiding method involves first, finding the zeroth moment of the object
as a first order estimate, and then iteratively correcting the centroid by computing
the zeroth order moment of the object weighted by a Gaussian window function,
until the correction falls below a particular threshold value.
The width of the Gaussian window function is set by the object's half light radius.

@@

In this paper, we propose to find the centroid of stars by smoothing the image
of stars by a Gaussian kernel, and fitting a second order polynomial to
the $3\times3$ pixels around the brightest pixel of smoothed image.
We focus on studying the accuracy of this model by comparing the centroid
measurement errors with the ones obtained from the quartic approximate method used in SDSS.
In order to do so, we apply these methods to a large number of simulated
stars with different signal-to-noise ratio and size realizations. 
Error from star centroiding methods, will always have 
a theoretically set lower bound, known as the Cramer-Rao lower bound
(hereafter denoted by CRLB) which has an inverse relation with the
signal-to-noise-ratio of stars. Assuming the exact model of the PSF,
one can compute the CRLB on ther centroiding error as a function of the
parameters of the model (\eg FWHM).

This paper is structured as follows. In section \ref{sec:model}, we give a brief overview of the methods used in our investigation. In section \ref{sec:data}, we discuss the simulated data. In section \ref{sec:results}, we compare the results of our centroid measurement with those of SDSS method, once for a fixed set of full width at half maxima, and once for a set of fixed signal-to-noise-ratios. Finally, we discuss and conclude in \ref{sec:discussion}.

>>>>

\documentclass[12pt, preprint]{aastex}
\usepackage{graphicx}	% For figures
\usepackage{natbib}	% For citep and citep
\usepackage{amsmath}	% for \iint
\usepackage{bbm}
\usepackage[breaklinks]{hyperref}	% for blackboard bold numbers
\usepackage{hyperref}
\hypersetup{colorlinks}
\usepackage{color}
\usepackage{morefloats}
\definecolor{darkred}{rgb}{0.5,0,0}
\definecolor{darkgreen}{rgb}{0,0.5,0}
\definecolor{darkblue}{rgb}{0,0,0.5}
\hypersetup{ colorlinks,
linkcolor=darkblue,
filecolor=darkgreen,
urlcolor=darkred,
citecolor=darkblue }

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}

\begin{document}

\author{
  Mohammadjavad~Vakili\altaffilmark{1},
  David~W.~Hogg\altaffilmark{1,2,3}
\altaffiltext{1}{Center for Cosmology and Particle Physics, New York University}
\altaffiltext{2}{Center for Data Science, New York University}
\altaffiltext{3}{Max-Planck-Institut f\"ur Astronomie}
}

\title{Towards fast and optimal centroiding of faint stars}

@@

We present a modification of the 3$\times$3 polynomial method for estimating the centroid 
of stars in astronomical images. After simulating a large number of stars with relatively low 
signal-to-noise-ratio, and in a wide range of full width at half maxima, we compare the results
of our method to those obtained from the quartic approximation method used in SDSS photometric pipeline, 
as well as finding the centroid by fitting the exact PSF model which was used to generate the 
simulated data. Comparison of the root-mean-squared error in centroid measurement, resulting 
from these techniques to the Cramer-Rao lower bound of centroid estimates, 
suggests that, the modified 3$\times$3 polynomial trick, though not as accurate as PSF fitting method
, is able to deliver results sufficiently close to the Cramer-Rao lower bound.

@@

A common practice in astronomy is taking imaging data, and then finding the coordinates
of various light sources across the sky. Finding accuarate estimate of the center of point
sources, convolved with telescope point spread function (and atmospheric PSF in case of
ground based telescopes), and also pixel response function, is crucial to further steps of
astronomical image processing. For instance, proper measurement of the shapes of galaxies
requires interpolating the PSF from the position of sparsely located stars across the
image to the positions of galaxies. The accuracy of the PSF estimation therefore,
relies on how accurately we know the centroid of stars.

@@

Reaching the optimal measurement of the centroid of stars however, is limited
by the lack of knowledge about the exact shape of the PSF, and also presence of noise;
both sky noise and CCD readout noise. Thus, of particular interest is devising a fast method that returns
optimal estimates of centroids in a realistic range of signal-to-noise ratios.

To date, a number of softwares have been designed for the purpose of extracting astronomical
sources and making catalogs. One of these softwares is SExtractor \citep{sex},
whose centroiding method involves first, finding the zeroth moment of the object
as a first-order estimate, and then iteratively correcting the centroid by computing
the zeroth order moment of the object weighted by a Gaussian window function,
until the correction falls below a particular threshold value.
The width of the Gaussian window function is set by the object's half light radius.

@@

In this paper, we propose to find the centroid of stars by smoothing the image
of stars by a Gaussian kernel, and fitting a second order polynomial to
the $3\times$3 pixels around the brightest pixel of smoothed image.
We compare the accuracy of this method with, the naive 3\times$3 second-order 
polynomial method, the 3$\times$3 quartic approximation
method used in SDSS photometric pipeline, and fitting a PSF model to the 
star.

In order to do so, we apply these methods to a large number of simulated
stars with different signal-to-noise ratio and size realizations. 
Error from star centroiding methods, will always have 
a theoretically-set lower bound, known as the Cramer-Rao lower bound
(hereafter denoted by CRLB) which has an inverse relation with the
signal-to-noise-ratio of stars.

Given the analytic model for the PSF,
we derive an expression for CRLB on the centroiding error as
a function of the parameters of the PSF model (\eg FWHM),
and signal-to-noise-ratio of stars. We create two sets of simulations for which we 
know the CRLB, one with variable (constant) signal-to-noise ratio (FWHM), and one 
with variable (constant) FWHM (signal-to-noise ratio). After applying
different centroiding methods to the simulations,
we investigate how close these methods can get to saturating the CRLB for
various ranges of background noise and FWHM.

This paper is structured as follows. In section \ref{sec:method},
we give a brief overview of the centroiding methods used in our investigation.
In section \ref{sec:data}, we discuss the simulated data, and the CRLB on centroiding error.
In section \ref{sec:results}, we compare the results of our proposed method with 
those obtained from centroiding methods discussed in \ref{sec:method}, and 
with the CRLB derived in \ref{sec:data}. Finally, we discuss and conclude
in \ref{sec:discussion}.
